\chapter{Related Works}
\label{chap:relwork}

Currently there are several other middleware projects similar to DIRAC. The DFC was developed by the LHCb experiment collaboration which, however, has similar needs as other LHC experiments. In particular the ATLAS
Distributed Data Management system and the AliEn file catalog are the closest comparable solutions.


\section{ATLAS DDM}
The ATLAS experiment at LHC is a general-purpose particle detector designed to investigate physics at the energy
frontier. The output data rate, already reduced by the online trigger is  200-400Hz. Even with this reduction 
ATLAS records a huge amount of data – more than 10PB of raw collision data have been taken since the LHC
started running in November 2009. After the first pass of processing was done at CERN, the data is registered in 
the ATLAS Distributed Data Management system (DDM)\cite{ATLASDDM1}.	

Although a basic unit of data in the ATLAS DDM is a file, the basic operation unit is a dataset: they may be
transferred to the Grid sites, whereas single files may not. There is also one more layer of aggregation called 
the container, which is a collection of datasets. Datasets may overlap and in practice they do so in a 
hierarchical manner: production will use small datasets, referring to a few jobs processed at a site. These 
datasets are then aggregated into the main dataset that refers to a particular task in the ATLAS
production system. These main datasets then may be added to a container, where the output of several 
similar tasks is gathered. Datasets may be open (new files can be added to them), closed (no new files
may be added, but can have versions which differ in their content) and frozen (no new files may be added 
and no new versions created). 

The responsibilities of the ATLAS DDM cover data registration, transfers between sites, file deletion, dataset 
consistency ensuring (manage file loses on sites), enforcing ATLAS Computing model policies and monitoring. The
current implementation is called Rucio (the previous one was DQ2).


\subsection{Rucio}
In the current implementation files, datasets and containers follow an identical naming scheme which is composed 
of the scope and a name, called a data identifier~\cite{ATLAS-Rucio}. Scopes are used to separate production and 
individual users. Metadata associated with a data identifier is represented using attributes (e.g. for a file its 
availability, for a dataset whether it is open, closed or frozen,...) which are represented as key-value pairs. 
The set of available attributes is restricted. Some metadata attributes can be set by a user, e.g. physics 
attributes (run number). Metadata not set by users include system attributes (size, creation time,...). 
For datasets and containers it is possible that the value of metadata is a function of the metadata of its
constituents, e.g. the total size is the sum of the sizes of the constituents.

The Rucio Storage Element (RSE) is a repository for physical files. It is the smallest unit of
storage space addressable within Rucio. It has a unique identifier and a set of properties (e.g. supported 
protocols, storage type, physical space,...). Physical files stored on RSEs are identified by their Physical File 
Name (PFN).
The PFN is a fully qualified path identifying a replica of a file. The mapping between the file identifier and
the PFN is a function of the identifier, RSE and protocol. Replica management in Rucio is based on
replication rules defined on data identifier sets. A replication rule is owned by an account and defines the 
minimum number of replicas to be available on a list of RSEs.

\section{AliEn}

AliEn (ALICE Environment)~\cite{AliEn2} is a Grid framework developed by the ALICE
Collaboration and used in production for more than 10 years. 

The AliEn File Catalog is one of the key components of the AliEn system. It
provides the mapping between Logical File Names (LFNs) visible to the end user
and one or more Physical File Names (PFNs) which describe the physical location
of the file by identifying the name of a storage element and the path to the
local file. The LFNs are manipulated by users, one LFN can point to more PFNs.
To prevent duplicate entries, each LFN is associated with a Globally Unique
IDentifier (GUID).

The interface of the catalog is a hierarchical structure of directories and
files, similar to the UNIX file system. The directories and files in the File
Catalog have UNIX-like privileges.

It also extends the familiar file system paradigm to include information about
running processes in the system (in analogy to the \texttt{/proc} directory in Linux
systems). Each job inserted into AliEn Task Queue gets a unique id and a
corresponding \texttt{/proc/id} directory, where it can register temporary files,
standard input and output as well as all job products~\cite{AliEn1}.

The File Catalog is designed to allow each directory node in the hierarchy to
be supported by different database engines, possibly running on different hosts
and even having different internal table structures optimized for a particular
directory branch. 

From the user point of view, the catalog behaves as a single entity. Internally
it is divided between the LFN and GUID catalogs that are kept in separate
databases. Every LFN database has an index table used to locate the files and
a table that contains all the information about the concrete LFN (owner, group,
size,...).

In addition, there can be user-defined tables containing metadata.
AliEn supports file collections.  A file collection is a user-defined list of
entries that can be either other collections, LFNs or GUIDs. The collection
mimics a file in the file catalog, its size is the sum of sizes of
all its components. If a command is executed over a collection, it will be
executed on each entry of the collection (e.g. if a user issues a command to
replicate a collection, all the files of that collection will be replicated).

\section{Comparison}

The ATLAS DDM is made to be used by the ATLAS collaboration only, so it does not have to cover all the 
requests from other experiments, e.g. user-defined metadata. The DDM also introduces more layers of data 
structures above files: there are datasets, nested datasets and containers. Furthermore, 
there is no structure in the names of datasets and containers, they have just unique identifiers. 

AliEn is opened for other virtual organizations, however there are only 2 other virtual organizations using 
AliEn services, while there are more than 15 ones using DIRAC. Since the administration of the whole middleware 
solution requires high effort, a proposition was made to integrate smaller independent DIRAC installations into 
a single service to optimize operational costs. Thus the France-Grilles DIRAC service, 
after an initial pre-production period of 6 months, was put into production in May 2012~\cite{FGDIRAC}.
This enabled small user groups not capable of installing, configuring and maintaining  
such complex systems as DIRAC, to use it without the need to understand and operate the complex software and
configurations. From the data management point of view, the  structure is similar: there are files and datasets in a 
directory structured space, user can define new metadata.


Unlike DIRAC, AliEn uses the directory structure not solely for keeping the information about the files, 
but also introduces the \texttt{/proc} directory for temporary job files and outputs.